---
title: "Vehicle Fraud Detection"
author: "Liam Knight"
format:
  html:
    toc: true
    number-sections: true
---

# Descriptive analysis of dataset

```{r, include = FALSE}

# load packages
packages <- c('tidyverse', 'knitr')
lapply(packages, library, character.only=TRUE)

# read in data
df <- read.csv("data/fraud_oracle.csv", 
                header = TRUE,
                stringsAsFactors = TRUE,
                check.names = FALSE) 
```

```{r} 
# initial check for missing values
sum(sapply(df, function(x) sum(is.na(x))))
```

## Cleaning and inspection

Though there are no missing values in this dataset, there are a few errors and inconsistencies we need to fix: some misspellings/typos, inconsistency in variable naming conventions, a meaningless variable `PolicyNumber`, and two variables `PoliceReportFiled` and `WitnessPresent` that are coded as character strings, but should function as Boolean types.

```{r}
# clean up data
df <- df |>
  # fix typos and standardize variable name conventions
  mutate(Make = recode(Make, 
                       "Accura" = "Acura",
                       "Mecedes" = "Mercedes",
                       "Nisson" = "Nissan",
                       "Porche" = "Porsche")) |>
  rename(NumberOfSupplements = NumberOfSuppliments,
         AddressChangeClaim = AddressChange_Claim,
         DaysPolicyClaim = Days_Policy_Claim,
         DaysPolicyAccident = Days_Policy_Accident,
         FraudFoundP = FraudFound_P) |>
  # PolicyNumber doesn't add any meaningful information, remove it
  select(!PolicyNumber) |>
  # PoliceReportFiled and WitnessPresent are really Booleans
  mutate(PoliceReportFiled = as.logical(recode(PoliceReportFiled, 
                                               'No' = 0, 'Yes' = 1)),
        WitnessPresent = as.logical(recode(WitnessPresent, 
                                           'No' = 0, 'Yes' = 1))) 
```

Now we inspect a few relevant features of the dataset.

```{r}
# function to get descriptive info about each variable
variable_info <- function(df) {
  n_rows <- nrow(df)
  tab <- tibble(
    variable = names(df),
    class = map_chr(df, ~ class(.x)),
    n_unique = map_int(df, ~ n_distinct(.x)),
    # run a custom function on each column to get character vector of 
    # values for each column that appear in descending order of counts
    values_descending = map_chr(df, ~ {
      tab <- sort(table(as.character(.x)), decreasing = TRUE)
      if (length(tab) == 0) return(NA_character_)
      paste0(names(tab), collapse = ", ")
    })
  )
  return(tab)
}

kable(variable_info(df))

```
## Odd values

Though we had already confirmed there are no missing values in this dataset, a few columns have some unexpected values of 0: `DayOfWeekClaimed`, `MonthClaimed`, `Age`. Note that as the variable `Age` refers to the age of the policyholder, 0 does not make sense as a value for any of these variables.

```{r}
# table that indicates how many times a value of 0 or "0" appears in each column
tab_zeroes <- df |>
  select(all_of(c("DayOfWeekClaimed", "MonthClaimed", "Age"))) |>
  summarise(across(everything(), ~ sum(as.character(.) == "0"))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "count")

kable(tab_zeroes)

```
Most of the values of 0 are in the `Age` column. When considering how to treat these values that are effectively missing values, unfortunately we are not able to speak to the data owner or source to understand where these values might have come from, which could help inform the decision about how to handle them. Additionally, there is no obvious reasonable method of imputation here. Thus, since 320 values comprise only 2.08% of the entire dataset of 15420 values and therefore a very small proportion, we will drop the rows that have these values of 0.

```{r}
df <- df |>
  filter(!if_any(all_of(c("DayOfWeekClaimed", "MonthClaimed", "Age")), 
                 ~ coalesce(as.character(.) == "0")))
```

Now we can perform some exploratory data analysis.

# Exploratory Data Analysis (EDA)

```{r}
# 




```



